\newglossaryentry{node}
{
        name=node,
        description={TODO! A fundamental entity in the Golem p2p network.}
}

\newglossaryentry{comp}
{
        name=computational resources,
        description={TODO! Any resources which allow to run a program. 
					In practice a single traditional computer or a cluster of computers.}
}

\newglossaryentry{requestor}
{
        name=requestor,
        description={TODO! A node which needs computational power. It defines the payload descriptor 
					and sends out demand order to find requestors.}
}

\newglossaryentry{provider}
{
        name=provider,
        description={TODO! A node that has computation resources to offer. It creates and sends out the offers. 
					One provider might be managing many computers (e.g. in case of a datacenter)}
}

\newglossaryentry{payload}
{
        name=payload,
        description={TODO! The computer program code, Docker or VM image, as well as the
					configuration and/or input data, which the requestor intends to have deployed
					and executed by the provider node(s)}
}

\newglossaryentry{plane}
{
        name=plane,
        description={}
}

\newglossaryentry{layer}
{
        name=layer,
        description={}
}

\newglossaryentry{namespace}
{
        name=namespace,
        description={}
}

\newglossaryentry{aspDim}
{
        name=aspect dimension,
        description={}
}

\newglossaryentry{service}
{
        name=service,
        description={}
}

\newglossaryentry{resource}
{
        name=resource,
        description={}
}

\newglossaryentry{demand}
{
        name=demand,
        description={TODO! Demand is created by Requestors and contains description of the
					payload, including properties telling what capabilities the payload might
					require (e.g. specific hardware architecture or features) and optionally other
					conditions which ought to be satisfied by the provider nodes (e.g. asking only
					for nodes which are located within specific country). Most of the content of the
					Demand is opaque for the Golem network}
}

\newglossaryentry{offer}
{
        name=offer,
        description={TODO! Offer is created by Provider nodes and contains the description of the
					computing resources the node has to offer, as well as any optional conditions
					for their use (e.g. limiting their availability only to certain requestor
					nodes).}
}

\newglossaryentry{proposal}
{
        name=proposal,
        description={TODO! The object which includes all the Demand/Offer/Proposal properties. 
					This is a JSON object in "flat convention" - where keys are full property names and 
					their values indicate properties.}
}

\newglossaryentry{agreement}
{
        name=agreement,
        description={TODO! The object which includes all the Demand/Offer/Proposal properties. 
					This is a JSON object in "flat convention" - where keys are full property names and 
					their values indicate properties.}
}

\newglossaryentry{invoice}
{
        name=invoice,
        description={TODO! An Invoice is an artifact issued by the Provider to the Requestor, 
					in the context of a specific Agreement. It indicates the total Amount owed by 
					the Requestor in this Agreement. 
					No further Debit Notes shall be issued after the Invoice is issued. 
					The issue of Invoice signals the Termination of the Agreement. 
					No Activity execution is allowed after the Invoice is issued.}
}

\newglossaryentry{property}
{
        name=property,
        description={}
}

\newglossaryentry{debit}
{
        name=debit note,
        description={A Debit Note is an artifact issued by the Provider to the Requestor, in the context of a specific Activity. 
					It is a notification of Total Amount Due incurred by the Activity until the moment the Debit Note is issued. 
					This is expected to be used as trigger for payment in upfront-payment or pay-as-you-go scenarios.}
}

\newacronym{gn}{GN}{Golem Node}
\newacronym{ga}{GA}{Golem Agent}
\newacronym{gc}{GC}{Golem Client}
\newacronym{gad}{GAD}{Golem Agent Daemon}
\newacronym{gac}{GAC}{Golem Agent Console}
\newacronym{gm}{GM}{Golem Manager}
\newacronym{gsm}{GSM}{Golem Schedule Manager}
\newacronym{gpg}{GPG}{Golem Payment Gateway}
\newacronym{gtk}{GTK}{Golem Toolkit}
\newacronym{gms}{GMS}{Golem Market Service}
\newacronym{gas}{GAS}{Golem Activity Service}
\newacronym{gps}{GPS}{Golem Payment Service}
\newacronym{gns}{GNS}{Golem Net Service}
\newacronym{gsb}{GSB}{Golem Service Bus}
\newacronym{gmsapi}{GMS API}{Golem Market Service API}
\newacronym{gasapi}{GAS API}{Golem Activity Service API}
\newacronym{gpsapi}{GPS API}{Golem Payment Service API}
\newacronym{gar}{GAR}{Golem Abstraction Resource}
\newacronym{res}{RES}{Resource}
\newacronym{up}{UP}{User Plane}
\newacronym{epp}{EPP}{External Provider Plane}
\newacronym{gp}{GP}{Golem Plane}
\newacronym{gacp}{Gacp}{Golem agent console point}
\newacronym{grp}{GRp}{Golem REST point}
\newacronym{gmp}{Gmp}{Golem market point}
\newacronym{gap}{Gap}{Golem activity point}
\newacronym{gpp}{Gpp}{Golem payment point}
\newacronym{gnp}{Gnp}{Golem net point}
\newacronym{Grp}{Grp}{Golem resource point}
\newacronym{Gsp}{Gsp}{Golem service point}
\newacronym{gntw}{Gntw}{Golem network}
\newacronym{pl}{PL}{Platform Layer}
\newacronym{il}{IL}{Integration Layer}
\newacronym{bl}{BL}{Business Layer}
